func-descr.template

### 2025-11-04 reports/capstone-movielens.site/05.appendix.Rmd

::: { #func.open_logfile.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_start_date.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_start_date.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_end_date.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_end_date.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_log.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_log.Usage .sidebar }
##### Usage
\

@@

::: { #func.get_log1.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_log1.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_log1.Usage .sidebar }
##### Usage
\

@@

::: { #func.get_log2.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_log2.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_log2.Usage .sidebar }
##### Usage
\

@@

::: { #func.get_log3.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_log3.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_log3.Usage .sidebar }
##### Usage
\

@@

::: { #func.get_log4.Usage .sidebar }
##### Usage
\

@@

::: { #func.print_log4.Usage .sidebar }
##### Usage
\

@@

::: { #func.put_log4.Usage .sidebar }
##### Usage
\

@@

::: { #func.clamp.Usage .sidebar }
##### Usage
\

@@

::: { #func.msg.set_arg.Usage .sidebar }
##### Usage
\

@@

::: { #func.msg.glue.Usage .sidebar }
##### Usage
\

@@

::: { #func.mse.Usage .sidebar }
##### Usage
\

@@

::: { #func.mse_cv.Usage .sidebar }
##### Usage
\

@@

::: { #func.rmse.Usage .sidebar }
##### Usage
\

@@

::: { #func.rmse2.Usage .sidebar }
##### Usage
\

@@

::: { #func.make_source_datasets.Usage .sidebar }
##### Usage
\

@@

This function produces datasets listed in the [Value](#func.make_source_datasets.value) section below and described in detail in [Appendix B: Model Training Datasets] using the `edx` as the original dataset.

@@

The pseudocode of the [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is shown below:
```{r eval=FALSE}
make_source_datasets <- function(){

@@

  put_log("Function: `make_source_datasets`: Dataset created: movie_map")

@@

  str(date_days_map)
  put_log("Function: `make_source_datasets`: Dataset created: date_days_map")

@@

  #> We will use K-fold cross validation as explained in 
  #> Section 29.6.1: "K-fold validation" of the Cource Textbook:
  #> https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation
  edx_CV <- lapply(kfold_index, function(fold_i){...}

@@

::: {.noteblock data-latex=""}
The complete source code of the [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is available in the [Initializing input datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L88) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) script on _GitHub_.
:::
:::

@@

@@

::: { #func.sample_train_validation_sets.Usage .sidebar }
##### Usage
\

@@

::: { #func.sample_train_validation_sets.args.data .sidebar }
  - **data:** ;

:::
::: { #func.sample_train_validation_sets.args.seed .sidebar }
  - **seed:** ;

@@

### Model Training Functions

@@

::: { #func.union_cv_results.Usage .sidebar }
##### Usage
\

@@

::: { #func.get_best_param.result.Usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.Usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.left.n.Usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.left_detailed.Usage .sidebar }
##### Usage
\

@@

## Model Training: Support Functions
\

@@

::: { #func.@@func.Usage .sidebar }
##### Usage
\

@@

# Appendix B: Model Training Datasets {#appndx_b}
\

@@

Note that we use the `splitGenreRows` function to split rows of the original dataset:
```{r eval=FALSE}
splitGenreRows <- function(data){
  put("Splitting dataset rows related to multiple genres...")
  start <- put_start_date()
  gs_splitted <- data |>
    separate_rows(genres, sep = "\\|")
  put("Dataset rows related to multiple genres have been splitted to have single genre per row.")
  put_end_date(start)
  gs_splitted
}
```

>>>>

::: { #func.open_logfile.usage .sidebar }
##### Usage
\

@@

::: { #func.print_start_date.usage .sidebar }
##### Usage
\

@@

::: { #func.put_start_date.usage .sidebar }
##### Usage
\

@@

::: { #func.print_end_date.usage .sidebar }
##### Usage
\

@@

::: { #func.put_end_date.usage .sidebar }
##### Usage
\

@@

::: { #func.print_log.usage .sidebar }
##### Usage
\

@@

::: { #func.put_log.usage .sidebar }
##### Usage
\

@@

::: { #func.get_log1.usage .sidebar }
##### Usage
\

@@

::: { #func.print_log1.usage .sidebar }
##### Usage
\

@@

::: { #func.put_log1.usage .sidebar }
##### Usage
\

@@

::: { #func.get_log2.usage .sidebar }
##### Usage
\

@@

::: { #func.print_log2.usage .sidebar }
##### Usage
\

@@

::: { #func.put_log2.usage .sidebar }
##### Usage
\

@@

::: { #func.get_log3.usage .sidebar }
##### Usage
\

@@

::: { #func.print_log3.usage .sidebar }
##### Usage
\

@@

::: { #func.put_log3.usage .sidebar }
##### Usage
\

@@

::: { #func.get_log4.usage .sidebar }
##### Usage
\

@@

::: { #func.print_log4.usage .sidebar }
##### Usage
\

@@

::: { #func.put_log4.usage .sidebar }
##### Usage
\

@@

::: { #func.clamp.usage .sidebar }
##### Usage
\

@@

::: { #func.msg.set_arg.usage .sidebar }
##### Usage
\

@@

::: { #func.msg.glue.usage .sidebar }
##### Usage
\

@@

::: { #func.mse.usage .sidebar }
##### Usage
\

@@

::: { #func.mse_cv.usage .sidebar }
##### Usage
\

@@

::: { #func.rmse.usage .sidebar }
##### Usage
\

@@

::: { #func.rmse2.usage .sidebar }
##### Usage
\

@@

::: { #func.make_source_datasets.usage .sidebar }
##### Usage
\

@@

::: { #func.make_source_datasets.details .sidebar }
##### Details
\

This function produces datasets listed in the [Value](#func.make_source_datasets.value) section below and described in detail in [Appendix B: Models Training Datasets] using the `edx` as the original dataset.

::: {.noteblock data-latex=""}
The function calls internally the following functions described below:

  - [splitGenreRows](#) to create the [`edx.sgr` Object];
  - [sample_train_validation_sets](#) to create *train* and *validation* sets of the inner datasets of the [`edx_CV` Object].

@@

The source code of the [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is shown below:
```{r eval=FALSE}
make_source_datasets <- function(){
  put_log("Function: `make_source_datasets`: Creating source datasets...")

@@

  put_log("Function: `make_source_datasets`: Dataset created: movie_map")
  put(summary(movie_map))

@@

  str(date_days_map)
  put_log("Function: `make_source_datasets`: Dataset created: date_days_map")
  put(summary(date_days_map))

@@

  #> We will use K-fold cross validation as explained in 
  #> Section 29.6.1: "K-fold validation" of the Cource Textbook:
  #> https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation
  #> We are going to compute the following version of the MSE introducing in that section:

  # $$
  #   \mbox{MSE}(\lambda) \approx\frac{1}{B} \sum_{b = 1}^B \frac{1}{N}\sum_{i = 1}^N \left(\hat{y}_i^b(\lambda) - y_i^b\right)^2 
  # $$

  start <- put_start_date()
  edx_CV <- lapply(kfold_index,  function(fold_i){

    # train_set <- mutate(train_set, userId = factor(userId), movieId = factor(movieId))
    # train.sgr <- mutate(train.sgr, userId = factor(userId), movieId = factor(movieId))

    list(train_set = train_set,
         train_mx = train_mx,
         train.sgr = train.sgr,
         validation_set = validation_set)
  })
  put_end_date(start)
  put_log("Function: `make_source_datasets`: 
Set of K-Fold Cross Validation datasets created: edx_CV")

  tuning_sets <- edx |>
    sample_train_validation_sets(5423)

@@

::: {.noteblock data-latex=""}
The source code of the [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is also available in the [Initializing input datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L88) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) script on _GitHub_.
:::
:::

#### [init_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) Function {#func.init_source_datasets}
\

::: { #func.init_source_datasets.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
movielens_datasets <- init_source_datasets()

This function produces datasets listed in the [Value](#func.init_source_datasets.value) section below and described in detail in [Appendix B: Models Training Datasets] using the `edx` as the original dataset.

::: { #func.init_source_datasets.value .sidebar }
##### Value
\

::: { #func.init_source_datasets.code .sidebar }
##### Source Code
\

The simplified source code of the [init_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is shown below:
```{r eval=FALSE}
init_source_datasets <- function(){
  put_log("Method `init_source_datasets`: 
Initializing sourse datasets...")

  if(file.exists(movielens_datasets_file_path)){
    movielens_datasets <- load_movielens_data_from_file(movielens_datasets_file_path)
  } else if(file.exists(movielens_datasets_zip)) {
    put_log("Method `init_source_datasets`: 
Unzipping MovieLens data file from zip-archive: {movielens_datasets_zip}...")

    unzip(movielens_datasets_zip, movielens_datasets_file_path)

    if(!file.exists(movielens_datasets_file_path)) {
      put_log("Method `init_source_datasets`: 
File does not exists: {movielens_datasets_file}.")
      stop("Failed to unzip MovieLens data zip-archive.")
    }

    movielens_datasets <- load_movielens_data_from_file(movielens_datasets_file_path)
  } else {
    put_log("Method `init_source_datasets`: 
Creating datasets...")
    library(edx.capstone.movielens.data)
    put_log("Method `init_source_datasets`: 
Library attached: 'edx.capstone.movielens.data'")

    movielens_datasets <- make_source_datasets()
    put("Method `init_source_datasets`: 
All required datasets have been created.")

    put_log("Method `init_source_datasets`: 
Saving newly created input datasets to file...")
    save(movielens_datasets, file =  movielens_datasets_file_path)

    if(!file.exists(movielens_datasets_file_path)) {
      put_log("Method `init_source_datasets`: 
File was not created: {movielens_datasets_file}.")
      warning("MovieLens data was not saved to file.")
    } else {
      put_log("Method `init_source_datasets`: 
Datasets have been saved to file: {movielens_datasets_file_path}.")

      put_log("Method `init_source_datasets`: 
Creating zip-archive: {movielens_datasets_zip}...")

      zip(movielens_datasets_zip, movielens_datasets_file_path)

      if(!file.exists(movielens_datasets_zip)){
        put_log("Method `init_source_datasets`: 
Failed to zip file: {movielens_datasets_file_path}.")
        warning("Failed to zip MovieLens data file.")
      } else {
        put_log("Method `init_source_datasets`: 
Zip-archive created: {movielens_datasets_zip}.")
      }
    }
  }
  movielens_datasets
}
```

::: {.noteblock data-latex=""}
The complete source code of the [init_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) function is available in the [Initializing input datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L88) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L89) script on _GitHub_.
:::
:::

@@

#### [load_movielens_data_from_file](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L4) Function {#func.load_movielens_data_from_file}
\

::: { #func.load_movielens_data_from_file.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
```
:::

::: { #func.load_movielens_data_from_file.args .sidebar }
##### Arguments
\

:::
::: { #func.load_movielens_data_from_file.args.seed .sidebar }
  - **file_path:** Data file full path.

::: { #func.load_movielens_data_from_file.details .sidebar }
##### Details
\

::: { #func.load_movielens_data_from_file.value .sidebar }
##### Value
\

::: { #func.load_movielens_data_from_file.code .sidebar }
##### Source Code
\

The source code of the [load_movielens_data_from_file](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L4) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [load_movielens_data_from_file](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L4) function is also available in the [Data processing functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L3) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L4) script on _GitHub_.
:::
:::

#### [splitByUser](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L14) Function {#func.splitByUser}
\

::: { #func.splitByUser.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
```
:::

::: { #func.splitByUser.args .sidebar }
##### Arguments
\

::: { #func.splitByUser.args.data .sidebar }
  - **data:** ;

::: { #func.splitByUser.details .sidebar }
##### Details
\

::: { #func.splitByUser.value .sidebar }
##### Value
\

::: { #func.splitByUser.code .sidebar }
##### Source Code
\

The source code of the [splitByUser](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L14) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [splitByUser](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L14) function is also available in the [Data processing functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L3) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L14) script on _GitHub_.
:::
:::

@@

Splits the input dataset passed in the [data](#func.sample_train_validation_sets.args.data) argument into *train* and *validation* datasets in the way described in [Section 31.6 *Cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#cross-validation) of the _Course Textbook (New Edition)_[@IDS2_31-6].

::: { #func.sample_train_validation_sets.usage .sidebar }
##### Usage
\

@@

::: { #func.sample_train_validation_sets.args.data .sidebar }
  - **data:** Input dataset to split;

:::
::: { #func.sample_train_validation_sets.args.seed .sidebar }
  - **seed:** Seed value;

@@

#### [mutateDateTimeAndDays](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L66) Function {#func.mutateDateTimeAndDays}
\

::: { #func.mutateDateTimeAndDays.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
```
:::

::: { #func.mutateDateTimeAndDays.args .sidebar }
##### Arguments
\

::: { #func.mutateDateTimeAndDays.args.data .sidebar }
  - **data:** ;

::: { #func.mutateDateTimeAndDays.details .sidebar }
##### Details
\

::: { #func.mutateDateTimeAndDays.value .sidebar }
##### Value
\

::: { #func.mutateDateTimeAndDays.code .sidebar }
##### Source Code
\

The source code of the [mutateDateTimeAndDays](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L66) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [mutateDateTimeAndDays](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L66) function is also available in the [Data processing functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L3) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L66) script on _GitHub_.
:::
:::

#### [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) Function {#func.splitGenreRows}
\
Splits rows of the original dataset passed in the [data](#) argument for movies that belong to multiple genres.

::: { #func.splitGenreRows.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
dat.sgr <- splitGenreRows(data)
```
:::

::: { #func.splitGenreRows.args .sidebar }
##### Arguments
\

::: { #func.splitGenreRows.args.data .sidebar }
  - **data:** A dataset containing rows corresponding to movies that belong to multiple genres..

::: { #func.splitGenreRows.details .sidebar }
##### Details
\

This function is used internally in the [make_source_datasets](#func.make_source_datasets) function described above in the [Initializing Input Datasets] section to create the `edx.sgr` dataset described below in the [`edx.sgr` Object] section of the [Appendix B: Models Training Datasets].

::: { #func.splitGenreRows.value .sidebar }
##### Value
\

::: { #func.splitGenreRows.code .sidebar }
##### Source Code
\

The source code of the [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) function is also available in the [Data processing functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L3) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) script on _GitHub_.
:::
:::

#### [filter_noMore_nratings](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L82) Function {#func.filter_noMore_nratings}
\

::: { #func.filter_noMore_nratings.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
```
:::

::: { #func.filter_noMore_nratings.args .sidebar }
##### Arguments
\

::: { #func.filter_noMore_nratings.args.data .sidebar }
  - **data:** ;

:::
::: { #func.filter_noMore_nratings.args.seed .sidebar }
  - **seed:** ;

::: { #func.filter_noMore_nratings.details .sidebar }
##### Details
\

::: { #func.filter_noMore_nratings.value .sidebar }
##### Value
\

::: { #func.filter_noMore_nratings.code .sidebar }
##### Source Code
\

The source code of the [filter_noMore_nratings](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L82) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [filter_noMore_nratings](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L82) function is also available in the [Data processing functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L3) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L82) script on _GitHub_.
:::
:::

### Models Training Functions

@@

::: { #func.union_cv_results.usage .sidebar }
##### Usage
\

@@

::: { #func.get_best_param.result.usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.left.n.usage .sidebar }
##### Usage
\

@@

::: { #func.data.plot.left_detailed.usage .sidebar }
##### Usage
\

@@

## Models Training: Support Functions
\

@@

::: { #func.@@func.usage .sidebar }
##### Usage
\

@@

# Appendix B: Models Training Datasets {#appndx_b}
\

@@

Note that we use the [splitGenreRows](#func.splitGenreRows) function described above to split rows of the original dataset:
```{r eval=FALSE}
splitGenreRows <- function(data){
  put("Splitting dataset rows related to multiple genres...")
  start <- put_start_date()
  gs_splitted <- data |>
    separate_rows(genres, sep = "\\|")
  put("Dataset rows related to multiple genres have been splitted to have single genre per row.")
  put_end_date(start)
  gs_splitted
}
```

### 2025-11-04 reports/capstone-movielens.site/work/func-descr.template.Rmd

::: { #func.log_func_tmpl.Usage .sidebar }
##### Usage
\

@@

::: { #func.helper-func.tmpl.Usage .sidebar }
##### Usage
\

@@

::: { #func.dat_func.tmpl.Usage .sidebar }
##### Usage
\

@@

::: { #func.dat_func.tmpl.args. .sidebar }
  - **:** ;

@@

::: { #func.datproc_func.tmpl.Usage .sidebar }
##### Usage
\

>>>>

::: { #func.log_func_tmpl.usage .sidebar }
##### Usage
\

@@

::: { #func.helper-func.tmpl.usage .sidebar }
##### Usage
\

@@

::: { #func.dat_func.tmpl.usage .sidebar }
##### Usage
\

@@

::: { #func.dat_func.tmpl.args.data .sidebar }
  - **data:** ;

@@

::: { #func.datproc_func.tmpl.usage .sidebar }
##### Usage
\
